% arara: pdflatex: { shell: yes }
\documentclass[a4paper,10pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{titlesec}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{xcolor}

% Set link color to classic blue and underline
\hypersetup{
    colorlinks=true,
    linkcolor=black,
    urlcolor=blue,
    citecolor=black,
    filecolor=black
}

% Formatting section titles
\titleformat{\section}{\large\bfseries}{}{0em}{}
\titleformat{\subsection}{\bfseries}{}{0em}{}

\begin{document}

\begin{center}
    {\LARGE \textbf{Erick Gonzalez Parada}} \\
    Puebla, Mexico $\cdot$ \underline{\href{https://linkedin.com/in/erick-gonzalez-810883335}{linkedin.com/in/erick-gonzalez-810883335}} \\
    +52 2228498841 $\cdot$ \underline{\href{mailto:erick.parada101@gmail.com}{erick.parada101@gmail.com}}
\end{center}

\section*{Summary}
I'm a software developer with expertise in full-stack development (both web and desktop). I love making my software available on all operating system platforms and have an interest in cybersecurity. Passionate about Linux, competitive programming, and chess. Currently a student at UDLAP (Universidad de las Américas Puebla), set to graduate in 2026.

\section*{Projects}
\subsection*{Courses Academy Video Platform (Full-Stack)}
\begin{itemize}
    \item Developed a platform where users can view courses created by administrators. Admins can add videos and manage other admins.
    \item Implemented CI/CD practices using Git and GitHub to improve efficiency, reducing bugs and delivery times.
    \item Designed user authentication flows on the frontend using JWT (e.g., login, signup, token storage, and renewal).
    \item Fully developed backend in JavaScript to avoid third-party services, allowing self-hosting and improved testability.
\end{itemize}

\subsection*{Fully Distributed GUI Application “idealCalculator” (Raylib)}
\begin{itemize}
    \item Developed \textit{idealCalculator}, a cross-platform numerical integration tool using C and Raylib.
    \item Implemented the Composite Trapezoidal and Midpoint Rules with seamless compatibility across Windows, macOS, and Linux.
\end{itemize}

\section*{Education}
\textbf{Universidad de las Américas Puebla (UDLAP)}\\
\textit{Computer Systems Engineering}\\
Recognition: Interim Board of Computer Systems Engineering; Event Coordinator\\
Cholula, Mexico | Graduate May 2026

\section*{Additional Skills}
\begin{itemize}
    \item \textbf{Languages:} Spanish (native), English (100\%), German (70\%)
    \item \textbf{Programming:} C/C++, Rust, Java, JavaScript/TypeScript, Nextjs, Sveltejs, SQL, OQL, XML, Bash \& Batch
    \item \textbf{Technologies:} Linux, GitHub, CI/CD, Web Development
    \item \textbf{Profiles:} \underline{\href{https://github.com/HugeErick}{GitHub}}, \underline{\href{https://portafolio-delta-wheat.vercel.app/}{Web Portfolio}}
\end{itemize}

\end{document}

